Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools

### 2024-03-25 paper.qmd
edits

Another motivation for cycle network planning at the European level is the Road Infrastructure Safety Management (RISM) directive (2008/96/EC), which requires member states to implement a road safety management system (RSMS) for all public roads.
Specifically, "Member States shall ensure that the ranking of high accident concentration sections and the network safety ranking are carried out" [@directiv2008].
Given that 'safety' in this context is usefully quantified as the number of people killed or seriously injured (KSI) per distance traveled, the directive requires estimation of distance traveled by mode, down to the road link level.
For active modes, about which there is a paucity of data compared with motorized modes, this is a major challenge.
Better data to inform road safety policies and interventions is a motivation for better estimates of 'baseline' levels of physical activity at high levels of geographic resolution [@tait2023].

National governments are increasingly acting on the evidence.
In Ireland, the Road Safety Authority (RSA) has set the target of halving the number of road traffic deaths and serious injuries by 2030 [@national2021].
Doing so while simultaneously enabling rapid uptake of active modes will require key travel corridors to be identified 'cycle proofed'.
Cycling in Ireland represents only 3% of total modal share as of the 2016 Census, but accounts for 20% of serious injuries and 7% of all fatalities, driving the development of cycling policy and infrastructure plans in Ireland.
The main frameworks underpinning these efforts are the Climate Action Plan [@climate2022a], the National Development Plan, and the National Roads 2040 strategy [@national2023].
<!-- Strategic alignment is also outlined by the Phoenix Park Transport and Mobility Options Study, which highlights public support for active travel interventions: while only 3% of trips in Ireland are made by cycling according to the latest Census data from 2016,^[ --> These policy frameworks are backed-up by public support for active travel interventions: while only 3% of trips in Ireland are made by cycling according to the latest Census data from 2016 "83% of respondents supported enhanced walking and cycling facilities" [@sustaina2019].

At the regional level within Ireland, the recently published Greater Dublin Area Transport Strategy [@greater] reinforces these findings of high support for and capability for cycling: "nearly a quarter of adults cycle at least once a week in the Dublin Metropolitan Area" with cycling in the Dublin area taking up to 60,000 cars off the road today.
Extrapolating this on a per population basis across Ireland, with around 40% of the population living in Dublin, suggests that around 150,000 cars could be removed nationwide just by achieving Dublin levels of cycling in all counties (notwithstanding existing cycling trips and differences in trip distances).
7 in 10 trips in Ireland are by car [@national] .
Cycling has the potential to replace a large proportion of these trips: "A high priority must also be given to cyclists, because trips by this mode have the potential to replace trips by private car, most specifically for short to medium distance trips, but increasingly for longer trips as ebikes extend the range of this mode" [@greater].

@@

> Direct routes with short journey times were found to be the most important positive variable for existing cyclists and non-cyclists in determining route choice.
> This is followed by infrastructure type, the number of junctions along the route, traffic speed and cyclist volumes.
> In terms if infrastructure, regardless of the level of cycling confidence, routes which have 'no facilities' or 'bus/cycle lanes' are the least favoured cycle route types.

@@

-   The results are *open access* meaning that any stakeholder in the planning system can access the evidence. This will help to democratize the transport planning process and make wider conversations about transport planning more evidence-based and less polarized [@lovelace2020], something that is particularly relevant given the potentially polarizing nature of pro-cycling interventions [@wild2017].
-   The results are fully reproducible (code to be released pending sign-off by TII's IT team), preventing 'cloud lock in' to a potentially monopolistic consultancy, and encouraging input from the wider open source community [@lovelace2021; @dhir2017].
-   By covering a wide range of trip purposes --- not just travel to work [@lovelace2017; @heinen2010] and travel school [@goodman2019] as covered in previously published research on open access national cycle network planning tools --- the results capture a higher proportion of cycling potential including key rural trips which are often under-represented in models of active travel.

>>>>

Another motivation for cycle network planning at the European level is the Road Infrastructure Safety Management (RISM) directive (2008/96/EC), which requires member states to implement a road safety management system (RSMS) for all public roads.
Specifically, "Member States shall ensure that the ranking of high accident concentration sections and the network safety ranking are carried out" [@directiv2008].
Given that 'safety' in this context is usefully quantified as the number of people killed or seriously injured (KSI) per distance traveled, the directive requires estimation of distance traveled by mode, down to the road link level.
For active modes, about which there is a paucity of data compared with motorized modes, this is a major challenge.
Better data to inform road safety policies and interventions is a motivation for better estimates of 'baseline' levels of physical activity at high geographic resolutions [@tait2023].

National governments are increasingly acting on the evidence.
In Ireland, the Road Safety Authority (RSA) has set the target of halving the number of road traffic deaths and serious injuries by 2030 [@national2021].
Doing so while simultaneously enabling rapid uptake of active modes will require key travel corridors to be identified and 'cycle proofed'.
Cycling in Ireland represents only 3% of total modal share as of the 2016 Census, but accounts for 20% of serious injuries and 7% of all fatalities.
Poor perception of safety has been found to represent the most important barrier to increased cycling in Ireland[@brick2018], and the need to improve road safety drives the development of national cycling policy and infrastructure plans.
The main frameworks underpinning these efforts are the Climate Action Plan [@climate2022a], the National Development Plan, and the National Roads 2040 strategy [@national2023].
<!-- Strategic alignment is also outlined by the Phoenix Park Transport and Mobility Options Study, which highlights public support for active travel interventions: while only 3% of trips in Ireland are made by cycling according to the latest Census data from 2016,^[ -->

At the regional level within Ireland, the recently published Greater Dublin Area Transport Strategy [@greater] reveals the high support for and capability for cycling: "nearly a quarter of adults cycle at least once a week in the Dublin Metropolitan Area" with cycling in the Dublin area taking up to 60,000 cars off the road today.
Extrapolating this on a per population basis across Ireland, with around 40% of the population living in Dublin, suggests that around 150,000 cars could be removed nationwide just by achieving Dublin levels of cycling in all counties (notwithstanding existing cycling trips and differences in trip distances).
7 in 10 trips in Ireland are by car [@national] .
Cycling has the potential to replace a large proportion of these trips: "A high priority must also be given to cyclists, because trips by this mode have the potential to replace trips by private car, most specifically for short to medium distance trips, but increasingly for longer trips as ebikes extend the range of this mode" [@greater].

@@

> Direct routes with short journey times were found to be the most important positive variable for existing cyclists and non-cyclists in determining route choice.
> This is followed by infrastructure type, the number of junctions along the route, traffic speed and cyclist volumes.
> In terms of infrastructure, regardless of the level of cycling confidence, routes which have 'no facilities' or 'bus/cycle lanes' are the least favoured cycle route types.

@@

-   The results are *open access* meaning that any stakeholder in the planning system can access the evidence. This will help to democratize the transport planning process and make wider conversations about transport planning more evidence-based and less polarized [@lovelace2020], something that is particularly relevant given the potentially polarizing nature of pro-cycling interventions [@wild2017].
-   The results are fully reproducible (code to be released pending sign-off by TII's IT team), preventing 'cloud lock in' to a potentially monopolistic consultancy, and encouraging input from the wider open source community [@lovelace2021; @dhir2017].
-   By covering a wide range of trip purposes --- not just travel to work [@lovelace2017; @heinen2010] and travel to school [@goodman2019] as covered in previously published research on open access national cycle network planning tools --- the results capture a higher proportion of cycling potential including key rural trips which are often under-represented in models of active travel.

### 2024-03-25 paper.tex
edits

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{European Transport Research Review}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools},
  pdfauthor={Robin Lovelace; Joey Talbot; Eugeni Vidal-Tortosa; Hussein Mahfouz; Elaine Brick; Peter Wright; Gary O'Toole; Dan Brennan; Suzanne Meade},
  pdfkeywords={Cycling, Open-source, Road Safety, Active Travel},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{keyword}
    Cycling \sep Open-source \sep Road Safety \sep 
    Active Travel
\end{keyword}
\end{frontmatter}

@@

\section{Introduction}\label{introduction}

@@

\centering{

@@

\end{figure}%

@@

\section{Tools for estimating cycling
potential}\label{tools-for-estimating-cycling-potential}

Tools for estimating geographical distribution of cycling potential have
advanced substantially in recent years. Developed by researchers and
developers in academic, public and private sectors, they have evolved
from simple area-based static models tailored to specific regions to
more complex and (in some cases) more generalizable tools. These tools
can be considered as a specific application of the concept of planning
support systems (PSS) , which are designed to help planners and
decision-makers to make more informed decisions \citep{geertman2009}.
Outputs include maps with results available at several levels of
analysis with levels of availability ranging from being only available
to researchers and in static maps to open access mapping systems
available to the public.

Previous studies have applied the concept of PSS to specific modes,
including walking \citep{bencekri2024}, public transport
\citep{barmentlo2012} and cycling \citep{bencekri2023}. While such prior
work provides insight into the potential for approaches to transport
planning that are both data-driven and participatory, the focus of this
paper is on tools that focus attention on the \emph{geographic
distribution} of cycling potential and which answer the question ``where
to build''. Not all papers reviewed in this section classify themselves
as, or even mention, PSS, but all of them can be considered as PSS. This
paper is informed by the thinking underlying PSS, including the idea
that the resulting tools are more effective if they are modular and can
inter-link with other tools and processes in the planning process: PSS
can be conceived as a ``toolbox with separate instruments that can work
and talk to each other effectively (as in the open systems concept) and
that will be selectively applied in varying configurations to support a
particular phase in the planning process'' \citep{geertman2002}. We
share this conception of PSS, emphasizing the benefits of a focus on a
particular mode (cycling) and a particular phase in the planning process
(strategic network development).

Some of the most relevant existing tools are summarised below and
presented in Table~\ref{tbl-tools}.

\begin{longtable}[t]{>{\raggedright\arraybackslash}p{13em}>{\raggedright\arraybackslash}p{15em}>{\raggedright\arraybackslash}p{12em}>{\raggedright\arraybackslash}p{5em}}

\caption{\label{tbl-tools}Summary of existing tools for estimating
cycling potential.}

\tabularnewline

\\
\toprule
Approach & Coverage and input data & Outputs and availability & License\\
\midrule
\cellcolor{gray!10}{Bicycle share model, Parkin et al. (2007)} & \cellcolor{gray!10}{England and Wales, Journeys to work OD census data, at the small-area (wards) level} & \cellcolor{gray!10}{Static tables in academic paper} & \cellcolor{gray!10}{Proprietary}\\
Analysis of  Cycling Potential (ACP), Transport for London (2010, 2016) & London, London Travel Demand Survey & Static tables, graphs, and heatmaps, not publicly available & Proprietary\\
\cellcolor{gray!10}{Prioritization index, Larsen et al. (2013)} & \cellcolor{gray!10}{Montreal, Survey, Road safety, and OD data} & \cellcolor{gray!10}{Map-based heatmap, not publicly available} & \cellcolor{gray!10}{Proprietary}\\
Usage intensity index, Zhang et al. (2014) & Belo Horizonte, Survey, census, and OD data & Static tables and maps, not publicly available & Proprietary\\
\cellcolor{gray!10}{The Cycling Potential Tool (CPT), Phillips and Range (2017)} & \cellcolor{gray!10}{Scotland, Environmental and socioeconomic data, at the small area (output areas) level} & \cellcolor{gray!10}{Maps showing cycling potential in each area} & \cellcolor{gray!10}{Proprietary}\\
\addlinespace
Propensity to Cycle Tool (PCT), Lovelace et al. (2017) & England and Wales, Journeys to work and school OD census data, at the small area (LSOA) level & Online maps, graphs, tables, publicly available at www.pct.bike & Open source\\
\cellcolor{gray!10}{The    Gross Potential for Cycling tool (CPC), Silva et al. (2021 and 2022)} & \cellcolor{gray!10}{21 Portuguese cities, Land use and socio-demographic data, at the small area (census tract) level} & \cellcolor{gray!10}{Static maps showing cycling potential in different areas} & \cellcolor{gray!10}{NA}\\
\bottomrule

\end{longtable}

@@

\section{Methods and data}\label{sec-methods}

\subsection{Disaggregation of origin-destination
data}\label{sec-disaggregation}

@@

\end{figure}%

\subsection{Additional trip purposes}\label{sec-trip-purposes}

@@

\end{figure}%

\subsection{Routing}\label{routing}

@@

\begin{minipage}{0.50\linewidth}
\includegraphics[width=2.60417in,height=\textheight]{images/naas_quietest_godutch.png}\end{minipage}%
%
\begin{minipage}{0.50\linewidth}
\includegraphics[width=2.60417in,height=\textheight]{images/naas_fastest_godutch.png}\end{minipage}%
\newline
\begin{minipage}{0.50\linewidth}
\includegraphics[width=2.60417in,height=\textheight]{images/newbridge_quietest_godutch.png}\end{minipage}%
%
\begin{minipage}{0.50\linewidth}
\includegraphics[width=2.60417in,height=\textheight]{images/newbridge_fastest_godutch.png}\end{minipage}%

@@

\end{figure}%

\subsection{Scenarios of cycling potential}\label{sec-scenarios}

@@

\end{figure}%

\subsubsection{Baseline}\label{baseline}

@@

\subsubsection{Near Market}\label{near-market}

@@

\subsubsection{Climate Action Plan}\label{climate-action-plan}

@@

\subsubsection{Go Dutch}\label{go-dutch}

@@

\subsubsection{Ebike}\label{ebike}

@@

\subsection{User interface}\label{sec-ui}

@@

\end{figure}%

\section{Results}\label{sec-results}

@@

\section{Discussion}\label{sec-discussion}

@@

\subsection{Uses of the tool in
practice}\label{uses-of-the-tool-in-practice}

@@

CRUSE is a ``great tool for getting started with project outline
documents''. It's an ``ideal tool for generating a strategic rationale
for estimating demand and building the business case in the absence of
cycle count data'' (Declan Keenan, Senior Transport Planner working on
active travel, personal communication). ``It's a good representation of
what's going on''

``I like to know that the data on the county maps is from POWSCAR data
which is gold standard''.

\subsection{Comparison with other tools and contribution to the
field}\label{comparison-with-other-tools-and-contribution-to-the-field}

@@

\section{Conclusions}\label{sec-conclusions}

@@

\section{List of abbreviations}\label{list-of-abbreviations}

@@

\section{Declarations}\label{declarations}

\subsection*{Availability of data and
material}\label{availability-of-data-and-material}
\addcontentsline{toc}{subsection}{Availability of data and material}

@@

\subsection*{Funding}\label{funding}
\addcontentsline{toc}{subsection}{Funding}

@@

\subsection*{Acknowledgements}\label{acknowledgements}
\addcontentsline{toc}{subsection}{Acknowledgements}

@@

  \bibliography{bibliography.bib,cruse.bib,references.bib}

>>>>

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\journal{European Transport Research Review}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools},
  pdfauthor={Robin Lovelace; Joey Talbot; Eugeni Vidal-Tortosa; Hussein Mahfouz; Elaine Brick; Peter Wright; Gary O'Toole; Dan Brennan; Suzanne Meade},
  pdfkeywords={Cycling, Open-source, Road Safety, Active Travel},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{keyword}
    Cycling \sep Open-source \sep Road Safety \sep 
    Active Travel
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, sharp corners, enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden]}{\end{tcolorbox}}\fi

@@

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

{\centering

@@

\end{figure}

@@

\hypertarget{tools-for-estimating-cycling-potential}{%
\section{Tools for estimating cycling
potential}\label{tools-for-estimating-cycling-potential}}

Tools for estimating geographical distribution of cycling potential have
advanced substantially in recent years. Developed by researchers and
developers in academic, public and private sectors, they have evolved
from simple area-based static models tailored to specific regions to
more complex and (in some cases) more generalizable tools. Outputs
include maps with results available at several levels of analysis with
levels of availability ranging from being only available to researchers
and in static maps to open access mapping systems available to the
public. Some of the most relevant existing tools are summarised below
and presented in Table~\ref{tbl-tools}.

\begin{table}

\caption{\label{tbl-tools}\textbf{?(caption)}}\begin{minipage}[t]{\linewidth}
\subcaption{\label{tbl-tools-1}Summary of existing tools for estimating cycling potential.}

{\centering

\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{13em}>{\raggedright\arraybackslash}p{15em}>{\raggedright\arraybackslash}p{12em}>{\raggedright\arraybackslash}p{5em}}
\toprule
Approach & Coverage and input data & Outputs and availability & License\\
\midrule
\cellcolor{gray!6}{Bicycle share model, Parkin et al. (2007)} & \cellcolor{gray!6}{England and Wales, Journeys to work OD census data, at the small-area (wards) level} & \cellcolor{gray!6}{Static tables in academic paper} & \cellcolor{gray!6}{Proprietary}\\
Analysis of  Cycling Potential (ACP), Transport for London (2010, 2016) & London, London Travel Demand Survey & Static tables, graphs, and heatmaps, not publicly available & Proprietary\\
\cellcolor{gray!6}{Prioritization index, Larsen et al. (2013)} & \cellcolor{gray!6}{Montreal, Survey, Road safety, and OD data} & \cellcolor{gray!6}{Map-based heatmap, not publicly available} & \cellcolor{gray!6}{Proprietary}\\
Usage intensity index, Zhang et al. (2014) & Belo Horizonte, Survey, census, and OD data & Static tables and maps, not publicly available & Proprietary\\
\cellcolor{gray!6}{The Cycling Potential Tool (CPT), Phillips and Range (2017)} & \cellcolor{gray!6}{Scotland, Environmental and socioeconomic data, at the small area (output areas) level} & \cellcolor{gray!6}{Maps showing cycling potential in each area} & \cellcolor{gray!6}{Proprietary}\\
\addlinespace
Propensity to Cycle Tool (PCT), Lovelace et al. (2017) & England and Wales, Journeys to work and school OD census data, at the small area (LSOA) level & Online maps, graphs, tables, publicly available at www.pct.bike & Open source\\
\cellcolor{gray!6}{The    Gross Potential for Cycling tool (CPC), Silva et al. (2021 and 2022)} & \cellcolor{gray!6}{21 Portuguese cities, Land use and socio-demographic data, at the small area (census tract) level} & \cellcolor{gray!6}{Static maps showing cycling potential in different areas} & \cellcolor{gray!6}{NA}\\
\bottomrule
\end{tabular}

\end{minipage}%

\end{table}

@@

\hypertarget{sec-methods}{%
\section{Methods and data}\label{sec-methods}}

\hypertarget{sec-disaggregation}{%
\subsection{Disaggregation of origin-destination
data}\label{sec-disaggregation}}

@@

\end{figure}

\hypertarget{sec-trip-purposes}{%
\subsection{Additional trip purposes}\label{sec-trip-purposes}}

@@

\end{figure}

\hypertarget{routing}{%
\subsection{Routing}\label{routing}}

@@

\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics[width=2.60417in,height=\textheight]{images/naas_quietest_godutch.png}

\end{minipage}%
%
\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics[width=2.60417in,height=\textheight]{images/naas_fastest_godutch.png}

\end{minipage}%
\newline
\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics[width=2.60417in,height=\textheight]{images/newbridge_quietest_godutch.png}

\end{minipage}%
%
\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics[width=2.60417in,height=\textheight]{images/newbridge_fastest_godutch.png}

\end{minipage}%

@@

\end{figure}

\hypertarget{sec-scenarios}{%
\subsection{Scenarios of cycling potential}\label{sec-scenarios}}

@@

\end{figure}

\hypertarget{baseline}{%
\subsubsection{Baseline}\label{baseline}}

@@

\hypertarget{near-market}{%
\subsubsection{Near Market}\label{near-market}}

@@

\hypertarget{climate-action-plan}{%
\subsubsection{Climate Action Plan}\label{climate-action-plan}}

@@

\hypertarget{go-dutch}{%
\subsubsection{Go Dutch}\label{go-dutch}}

@@

\hypertarget{ebike}{%
\subsubsection{Ebike}\label{ebike}}

@@

\hypertarget{sec-ui}{%
\subsection{User interface}\label{sec-ui}}

@@

\end{figure}

\hypertarget{sec-results}{%
\section{Results}\label{sec-results}}

@@

\hypertarget{sec-discussion}{%
\section{Discussion}\label{sec-discussion}}

@@

\hypertarget{uses-of-the-tool-in-practice}{%
\subsection{Uses of the tool in
practice}\label{uses-of-the-tool-in-practice}}

@@

\hypertarget{comparison-with-other-tools-and-contribution-to-the-field}{%
\subsection{Comparison with other tools and contribution to the
field}\label{comparison-with-other-tools-and-contribution-to-the-field}}

@@

\hypertarget{sec-conclusions}{%
\section{Conclusions}\label{sec-conclusions}}

@@

\hypertarget{list-of-abbreviations}{%
\section{List of abbreviations}\label{list-of-abbreviations}}

@@

\hypertarget{declarations}{%
\section{Declarations}\label{declarations}}

\hypertarget{availability-of-data-and-material}{%
\subsection*{Availability of data and
material}\label{availability-of-data-and-material}}
\addcontentsline{toc}{subsection}{Availability of data and material}

@@

\hypertarget{funding}{%
\subsection*{Funding}\label{funding}}
\addcontentsline{toc}{subsection}{Funding}

@@

\hypertarget{acknowledgements}{%
\subsection*{Acknowledgements}\label{acknowledgements}}
\addcontentsline{toc}{subsection}{Acknowledgements}

@@

  \bibliography{bibliography.bib,cruse.bib}
